### 2016-01-27 main text/draft0.0.tex
minor editing
Adding first bit of M101 data
Writing about ordering and tuning phase in the method section.

\newcommand \kpc        {\,{\rm kpc}}
\newcommand \sigmagas    {$\Sigma_{\rm \bld {gas }} $\ }
\newcommand \sigmatotalgas {$\Sigma_{\rm \bld {total\, gas }} $\ }
\newcommand \eqsigmagas    {\Sigma_{\rm \bld {gas }}}
\newcommand \sigmasfr     {$\Sigma_{\rm \bld {SFR }} $\ }
\newcommand \eqsigmasfr     {\Sigma_{\rm \bld {SFR }}}
\newcommand \sigmastar    {$\Sigma_{\rm \bld {star }} $\ }
\newcommand \eqsigmastar    {\Sigma_{\rm \bld {star }}}
\newcommand \halpha    {H$\alpha $\ }
\newcommand \um    {$\mu$m\ }
\newcommand \mice {$\mu$m}
\newcommand \nprime {N$^\prime$}
\newcommand \eqnprime {N^\prime}
\newcommand \Spitzer {{\it Spitzer }}
\newcommand \Galex {GALEX }
\newcommand \Herschel {{\it Herschel }}
\newcommand \aaj {A\&A}
\newcommand \aarv {A\&ARv}%: Astronomy and Astrophysics Review (the)
\newcommand \aas{A\&AS}%: Astronomy and Astrophysics Supplement Series
\newcommand \afz {Afz}%: Astrofizika
\newcommand \aj {AJ}%: Astronomical Journal (the)
\newcommand \apss {Ap\&SS}%: Astrophysics and Space Science
\newcommand \apj {ApJ}
\newcommand \apjs {ApJS}%: Astrophysical Journal Supplement Series (the)
\newcommand \araa {ARA\&A} %: Annual Review of Astronomy and Astrophysics
\newcommand \asp {ASP Conf. Ser.}%: Astronomy Society of the Pacific Conference Series
\newcommand \azh {Azh}%: Astronomicheskij Zhurnal
\newcommand \baas {BAAS}%: Bulletin of the American Astronomical Society
\newcommand \mem {Mem. RAS}%: Memoirs of the Royal Astronomical Society
\newcommand \mnassa {MNASSA}%: Monthly Notes of the Astronomical Society of Southern Africa
\newcommand \mnras {MNRAS} %: Monthly Notices of the Royal Astronomical Society
%\newcommand {Nature}%(do not abbreviate)
\newcommand \pasj {PASJ}%: Publications of the Astronomical Society of Japan
\newcommand \pasp {PASP}%: Publications of the Astronomical Society of the Pacific
\newcommand \qjras {QJRAS}%: Quarterly Journal of the Royal Astronomical Society
\newcommand \mex {Rev. Mex. Astron. Astrofis.}%: Revista Mexicana de Astronomia y Astrofisica
%\newcommand {Science }%}%(do not abbreviate)
\newcommand \sva {SvA}%: Soviet Astronomy
\newcommand \aap {APP} %:American Academy of Pediatrics
\newcommand \apjl {ApJL} %:The Astrophysical Journal Letters

@@

\title{Data Mining in Nearby Galaxies}
\author{rahmani.sahar }
\date{\today}
\author[S. Rahmani, et. al.]{S.~Rahmani$^{1}$\thanks{E-mail:
srahma49@uwo.ca}, H.~Teimoorinia$^{2}$, E.~Peeters$^{1}$, P.~Barmby$^{1}$\\
$^{1}$Department of Physics $\&$ Astronomy, Western University, London, ON N6A 3K7, Canada\\
$^{2}$Department of Physics $\&$ Astronomy, University of Victoria, Finnerty Road, Victoria, British Columbia, V8P 1A1, Canada}
\maketitle

@@

\begin{abstract} 
Here is the abstract.
%This is just first abstract for this paper!
%M31 was targeted by many probes, and there are variety of data available for this particular galaxy.
%M31 was mapped from X-ray to 21 cm emission. Also, spectroscopy data of different regions of M31 is available in various bands. 
%The vast availability of data for this galaxy makes it a suitable target for data mining and ﬁnding new patterns in this galaxy.
%In this project we are studying 10 regions in M31 from the bulge of the galaxy to the star forming ring. Using self organizing map (SOM) method, we classiﬁed these regions in different groups, based on total star formation rate, total stellar mass, total infrared emission and 3.6 \mice, 4.5 \mice, 5.8 \mice, 8 \mice, 24 \mice, 70 \um and 160 \mice, 2.6 mm, and 21 cm luminosity from the photometry data of the regions, and flux of 5.7 \mice, 6.2 \mice, 7.4 \mice, 7.6 \mice, 7.9\mice, 8.3 \mice, 8.6 \mice, 10.7 \mice, 11.23 \mice, 12.00 \mice, 12.62 \mice, 12.69 \mice, 14.0 \mice, 16.45 \mice, 17.04 \mice, PAH lines, [ArIII], [SIII], [SIV], [NeII], and [NeIII] as well as metallicity of these regions obtained from the spectroscopy data. 
%We divided these regions from 2 to 6 different classes, and for each class we are going to interpret results and check whether these classiﬁcations have a physical meaning or they are just statistically relate to each other. 
%As a result, we can predict some properties of regions in other galaxies by having only few data.

@@

\section*{ACKNOWLEDGMENTS}
S.R and P.B acknowledge research support from the Natural Sciences and Engineering Research Council of Canada and from the Academic Development Fund of the University of Western Ontario.
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%biblio
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\bibliographystyle{mnras}
\bibliography{ref_mining}

>>>>

\newcommand \kpc        {\,{\rm kpc}}
\newcommand \sigmagas    {$\Sigma_{\rm \bld {gas }} $\ }
\newcommand \sigmatotalgas {$\Sigma_{\rm \bld {total\, gas }} $\ }
\newcommand \eqsigmagas    {\Sigma_{\rm \bld {gas }}}
\newcommand \sigmasfr     {$\Sigma_{\rm \bld {SFR }} $\ }
\newcommand \eqsigmasfr     {\Sigma_{\rm \bld {SFR }}}
\newcommand \sigmastar    {$\Sigma_{\rm \bld {star }} $\ }
\newcommand \eqsigmastar    {\Sigma_{\rm \bld {star }}}
\newcommand \halpha    {H$\alpha $\ }
\newcommand \um    {$\mu$m\ }
\newcommand \mice {$\mu$m}
\newcommand \nprime {N$^\prime$}
\newcommand \boldit {\textbf{\textit{}}}
\newcommand \eqnprime {N^\prime}
\newcommand \Spitzer {{\it Spitzer }}
\newcommand \Galex {GALEX }
\newcommand \Herschel {{\it Herschel }}
\newcommand \aaj {A\&A}
\newcommand \aarv {A\&ARv}%: Astronomy and Astrophysics Review (the)
\newcommand \aas{A\&AS}%: Astronomy and Astrophysics Supplement Series
\newcommand \afz {Afz}%: Astrofizika
\newcommand \aj {AJ}%: Astronomical Journal (the)
\newcommand \apss {Ap\&SS}%: Astrophysics and Space Science
\newcommand \apj {ApJ}
\newcommand \apjs {ApJS}%: Astrophysical Journal Supplement Series (the)
\newcommand \araa {ARA\&A} %: Annual Review of Astronomy and Astrophysics
\newcommand \asp {ASP Conf. Ser.}%: Astronomy Society of the Pacific Conference Series
\newcommand \azh {Azh}%: Astronomicheskij Zhurnal
\newcommand \baas {BAAS}%: Bulletin of the American Astronomical Society
\newcommand \mem {Mem. RAS}%: Memoirs of the Royal Astronomical Society
\newcommand \mnassa {MNASSA}%: Monthly Notes of the Astronomical Society of Southern Africa
\newcommand \mnras {MNRAS} %: Monthly Notices of the Royal Astronomical Society
%\newcommand {Nature}%(do not abbreviate)
\newcommand \pasj {PASJ}%: Publications of the Astronomical Society of Japan
\newcommand \pasp {PASP}%: Publications of the Astronomical Society of the Pacific
\newcommand \qjras {QJRAS}%: Quarterly Journal of the Royal Astronomical Society
\newcommand \mex {Rev. Mex. Astron. Astrofis.}%: Revista Mexicana de Astronomia y Astrofisica
%\newcommand {Science }%}%(do not abbreviate)
\newcommand \sva {SvA}%: Soviet Astronomy
\newcommand \aap {APP} %:American Academy of Pediatrics
\newcommand \apjl {ApJL} %:The Astrophysical Journal Letters

@@

\title{Mining in archival data of Nearby Galaxies}
\author{rahmani.sahar }
\date{\today}
\author[S. Rahmani, et. al.]{S.~Rahmani$^{1}$\thanks{E-mail:
srahma49@uwo.ca}, H.~Teimoorinia$^{2}$, E.~Peeters$^{1}$, P.~Barmby$^{1}$\\
$^{1}$Department of Physics $\&$ Astronomy, Western University, London, ON N6A 3K7, Canada\\
$^{2}$Department of Physics $\&$ Astronomy, University of Victoria, Finnerty Road, Victoria, British Columbia, V8P 1A1, Canada}
\maketitle

@@

\begin{abstract} 
 The vast availability of data for nearby galaxies makes them  suitable target for data mining  in order to ﬁnd new patterns in these galaxies.
In this project we are studying data from 10 regions in M31 and 8 regions in M101.Based on available information for these two galaxies we classified these regions in different groups  using self organizing map (SOM) method.
%We divided these regions from 2 to 6 different classes, and for each class we are going to interpret results and check whether these classiﬁcations have a physical meaning or they are just statistically relate to each other. 
%As a result, we can predict some properties of regions in other galaxies by having only few data.

@@

\section*{ACKNOWLEDGMENTS}
S.R and P.B acknowledge research support from the Natural Sciences and Engineering Research Council of Canada and from the Academic Development Fund of the University of Western Ontario. This research has made use of the NASA/IPAC Extragalactic Database (NED), which is operated by the Jet Propulsion Laboratory, California Institute of Technology, under contract with the National Aeronautics and Space Administration.
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%biblio
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\bibliographystyle{mnras}
\bibliography{ref_mining.bib}

### 2016-01-27 sections/data0.0.tex
Writing about ordering and tuning phase in the method section.

     The 10 regions in M31 were chosen due to availability of the PAH data for them. 
     \cite{Dim15} studied PAH for those regions using {\it Spitzer}/Infrared Spectrograph\citep[IRS,][]{Houck04b} observations. 
     They used {\tiny PAHFIT IDL} tool~\citep{Smith07b} and measured fluxes and equivalent widths of PAH features as well as atomic line features.
     For spectroscopy part of our sample, we utilized their measured fluxes for PAH lines. 
     We used radiation hardness index instead of flux of individual atomic line, since they were only calculate the upper limit flux for most of the lines in most of the regions.

@@

%M101 data: observed ones;
    \subsection{M101 data}
     For M101 galaxy, we used PAH fluxes which had been measured using {\tiny PAHFIT IDL} tool and reported in \cite{Gordon08}. 
     We used data available on NED~\footnote{The NASA/IPAC Extragalactic Database (NED) is operated by the Jet Propulsion Laboratory, California Institute of Technology, under contract with the National Aeronautics and Space Administration.} website to gather our photometry data of M101. 
     We chose data from \GALEX FUV and NUV channels~\citep{depaz07}, IRAC channels 1 to 4, MIPS24 and 70~$\mu$m~\cite{Dale09}, and  PACS100 and 160~$\mu$m and SPIRE250, 350, and 500~$\mu$m~\cite{Kennicutt11}.
     Same as M31 data, we used $^{12}$CO (J:$1\rightarrow0$) line~\citep{Helfer03} and the atomic gas (H\,{\sc I}) 21~cm emission from \cite{Walter08} data.
     For SFR, stellar and total gas map we used followed method described in \cite{Rahmani16} to create these maps.

>>>>

     The 10 regions in M31 were chosen due to availability of the PAH data for them. 
     \cite{Dim15} studied PAH for those regions using {\it Spitzer}/Infrared Spectrograph\citep[IRS,][]{Houck04b} observations. 
     They used {\tiny PAHFIT IDL} tool~\citep{Smith07b} and measured fluxes and equivalent widths of PAH features as well as atomic line features.
     For spectroscopy part of our sample, we utilized their measured fluxes for PAH lines. 
     \cite{Dim15} also reported metallicity of these regions (using results from \cite{Sanders12} for most regions), which e used them in our input sample.
     We used radiation hardness index (RHI) instead of flux of individual atomic line, since they were only calculate the upper limit flux for most of the lines in most of the regions.

@@

%M101 data: observed ones;
    \subsection{M101 data}
     For M101 galaxy, we used PAH fluxes which had been measured using {\tiny PAHFIT IDL} tool and reported in \cite{Gordon08}.
     We also added their measurement of metallicity and RHI for these regions to our sample.
     We used data available on NED~\footnote{The NASA/IPAC Extragalactic Database (NED) is operated by the Jet Propulsion Laboratory, California Institute of Technology, under contract with the National Aeronautics and Space Administration.} website to gather our photometry data of M101. 
     We chose data from \GALEX FUV and NUV channels~\citep{depaz07}, IRAC channels 1 to 4, MIPS24 and 70~$\mu$m~\cite{Dale09}, and  PACS100 and 160~$\mu$m and SPIRE250, 350, and 500~$\mu$m~\cite{Kennicutt11}.
     Same as M31 data, we used $^{12}$CO (J:$1\rightarrow0$) line~\citep{Helfer03} and the atomic gas (H\,{\sc I}) 21~cm emission from \cite{Walter08} data.
     SFR(FUV + 24$\mu$m), stellar and total gas map were calculated followed method described in \cite{Rahmani16} to create these maps.

### 2016-01-27 sections/method0.0.tex
Writing about ordering and tuning phase in the method section.

 \subsubsection{Algorithm of SOM} 
 \label{sec: algorithm}

     Assuming we have a data set which contains vectors, \boldit{V} $\in \Re^n$ and we want to map them on S1 by S2 map. 
     We start by creating S1 $\times$ S2 empty neurons. 
     The initial arrangement of these neurons depends on a map's topology provided by user.
     Since the topology of the map does not have any affect on the final result, we chose hexagonal topology which is the default topology for SOMs.
     Then, we assign a random weight vector \boldit{W} $\in \Re^n$ to each node.
     The process of creating SOM, happens over series of $N$ iterations. 
     During each iteration the weight vectors might change according to the Kohonen learning rule (equation~\ref{equ: weight adj}). 
      In each iteration:
     \begin{enumerate}
     \item Choose a random vector from our data set.
     \item Calculate the euclidean distance for each node j as  $D_j^2= \sum_{i=0}^{i=n} (V_i - W_i)^2$, and find a neuron with ``$D_{j_{min}}$". This neuron is the winner node and is calling Best Matching Unit (BMU). 
     \item  Compute the radius of the neighbourhood of the BMU to find nodes within this radius. The weight vectors of these nodes will be affected in the next steps. This value is arbitrary and initially can be set to be as high as half of the SOM size and then it decades exponentially over each iteration:
   \begin{equation}
   r^t_{BMU} = r^0_{BMU}e^{(-t/\tau)}
   \end{equation}
   where $\tau$ is a decay constant and usually set to be the same as number of iterations, $N$. $r^0_{BMU}$ and $r^t_{BMU}$ is the radius of the neighbourhood at 0th and $t$th iteration, respectively. 
   \item Change the weight vectors of the BMU and all the nodes within r(t) as:
   \begin{equation}
   \label{equ: weight adj}
   w(t+1)=w(t)+L(t) \times R(t) \times(v(t)-w(t))
   \end{equation}
   where $L(t) = e^{(-t/\tau)}$ is the learning factor which prevents divergence of the SOM and $R(t)=exp(-\frac{D_j^2}{2r^t_{BMU}})$ is the influence rate. $R(t)$ determines how weight of nodes in the neighbourhood of BMU will change.
   \item  Repeat these steps for $N$ times.
    \end{enumerate}

    In order to create SOM, we used {\tiny MATLAB} neural network toolbox~\citep{sommatlab}. 
    We created different SOM with different sizes, initial neighbourhood, and iteration numbers to find the optimize result for our sample.
    We test our code by varying the number of iterations and find $N = 100$ is optimized for our sample.
    In Section~\ref{sec: test model} we used this toolbox to create SOMs from SED of galaxies and learn how this method works and how we can interpret the results.

>>>>

 \cite{Geach12} presented the application of the SOM and demonstrate the algorithm of the SOM in detail. In this section we are going to talk briefly about the algorithm of the SOM, how we create our maps and a test model which will help to interpret our results.

 \subsubsection{Algorithm of SOM} 
 \label{sec: algorithm}
     Assuming we have a data set which contains vectors, \boldit{V} $\in \Re^n$ and we want to map them on S1 by S2 map. 
     We start by creating S1 $\times$ S2 empty neurons. 
     The initial arrangement of these neurons depends on a map's topology provided by user.
     Since the topology of the map does not have any affect on the final result, we chose hexagonal topology which is the default topology for SOMs.
     Then, we assign a random weight vector \boldit{W} $\in \Re^n$ to each node.
     The process of creating SOM, happens over series of $N$ iterations. 
     During each iteration the weight vectors might change according to the Kohonen learning rule (equation~\ref{equ: weight adj}). 
      In each iteration:
     \begin{enumerate}
        \item Choose a random vector from our data set.
        \item Calculate the euclidean distance for each node j as  $D_j^2= \sum_{i=0}^{i=n} (V_i - W_i)^2$, and find a neuron with ``$D_{j_{min}}$". This neuron is the winner node and is calling Best Matching Unit (BMU). 
        \item  Compute the radius of the neighbourhood of the BMU to find nodes within this radius. The weight vectors of these nodes will be affected in the next steps. This value is arbitrary and initially can be set to be as high as half of the SOM size and then it decades exponentially over each iteration:
        \begin{equation}
            r^t_{BMU} = r^0_{BMU}e^{(-t/\tau)}
        \end{equation}
        where $\tau$ is a decay constant and usually set to be the same as number of iterations, $N$. $r^0_{BMU}$ and $r^t_{BMU}$ is the radius of the neighbourhood at 0th and $t$th iteration, respectively. 
        \item Change the weight vectors of the BMU and all the nodes within r(t) as:
        \begin{equation}
            \label{equ: weight adj}
            w(t+1)=w(t)+L(t) \times R(t) \times(v(t)-w(t))
        \end{equation}
        where $L(t) = L_0 e^{(-t/\tau)}$ is the learning factor which prevents divergence of the SOM and $R(t)=exp(-\frac{D_j^2}{2r^t_{BMU}})$ is the influence rate. $R(t)$ determines how weight of nodes in the neighbourhood of BMU will change.
        \item  Repeat these steps for $N$ times.
     \end{enumerate}

     In order to create SOM, we used {\tiny MATLAB} neural network toolbox~\citep[ NEWSOM,][]{matlabtolbox}.
     {\tiny MATLAB} code, works in two phases. Phase one is the ``ordering phase". 
     This phase starts with maximum neighbourhood distance (which is initially provided by user), and initial learning factor of 0.9 as a default values. 
     The ordering phase continues for requested number of iterations and during that steps changes the parameters in the way that in the last iteration learning factor is 0.02 and the neighbourhood distance is 1.

     The second phase is the ``tuning phase".
     In this phase the neighbourhood distance is at its minimum, but learning factor decreases very slowly.
     This minimum neighbourhood distance and slowly decreasing the leaning factor helps to fine tune the topology results and causes the more stable SOM. 
     The number of iterations in this tuning phase most be much more than the number iterations in ordering phase, to allow the tuning happens slowly. 
     We chose number of epochs the tuning phase be 3 times more than number of epochs in the ordering phase.

     We created different SOM with different sizes, initial neighbourhood, and iteration numbers to find the optimize result for our sample.
     We test our code by varying the number of iterations in ordering phase and find $N = 100$  is optimized for our sample.
     In Section~\ref{sec: test model} we used this toolbox to create SOMs from SED of galaxies and learn how this method works and how we can interpret the results.
